---
title: "adaboost"
output: html_document
---

```{r setup, include=FALSE}
# This chunk loads all the packages used in this homework
library(ISLR)
library(mlbench)
library(caret)
library(randomForest)
library(ranger)
library(gbm)
library(pdp)
library(pROC)
library(rpart)

library(ggplot2)
library(tidyverse)


# General figure set up
knitr::opts_chunk$set(
  # hide warning messages
  warning = FALSE
)
```

```{r input}
#source("preprocess.Rmd")
# adaboost
set.seed(2022)
ctrl <- trainControl(method = "repeatedcv",
                     classProbs = TRUE, 
                     summaryFunction = twoClassSummary)
```

```{r}

```



```{r adaboost}
library(parallel)
# Calculate the number of cores
num_cores <- detectCores() - 1
library(doParallel)
# create the cluster for caret to use
# CPU usage may go up to 100%
cl <- makePSOCKcluster(num_cores)
registerDoParallel(cl)

gbm.grid <- expand.grid(n.trees = c(2000, 3000, 4000, 5000, 6000),
                            interaction.depth = 1:8,
                            shrinkage = c(0.03,0.05,0.07),
                            n.minobsinnode = 1)
set.seed(2022)
model.gbm <- train(Category~.,
                   data = hcv.train.df,
                  tuneGrid = gbm.grid,
                  trControl = ctrl,
                  method = "gbm",
                  distribution = "adaboost",
                  metric = "Accuracy",
                  verbose = FALSE)

ggplot(model.gbm, highlight = TRUE)
model.gbm$bestTune
# Stop the parallel computing cluster
stopCluster(cl)
registerDoSEQ()
```

```{r svm_linear}
set.seed(2022)
model.svml <- train(Category ~ . , 
                  data = hcv.train.df, 
                  method = "svmLinear",
                  tuneGrid = data.frame(C = exp(seq(-4,5,len = 50))),
                  trControl = ctrl)

plot(model.svml, highlight = TRUE, xTrans = log)
model.svml$bestTune
```

```{r compare}
set.seed(2022)
model.ctree <- train(x = trainData.dummy,
                   y = hcv.train.df$Category,
                     method = "ctree",
                     tuneGrid = data.frame(mincriterion = 1 - exp(seq(-2, -1, leng = 100))),
                     metric = "Accuracy",
                     trControl = ctrl,
                     na.action = na.pass)
```

```{r save}
saveRDS(model.gbmA)
saveRDS(model.svml)
```


### Tuning parameters

#### Adaboost Model

We tested on different ranges of tuning parameters, including number of trees (`n.trees`), number of splits (`interaction.depth`) in each trees and the shrinkage parameter (`shrinkage`). We looked for the point where the best cross-validated ROC AUC is obtained. The best tuned parameters are number of trees = 4000, number of splits = 5, shrinkage = 0.002.

#### MARS Model

MARS model can take a wide degree of features and number of terms. For simplicity, we only consider the performance of MARS in the first four degrees and all terms. The best tuned parameter is degree = 3 and nprune = 6.

0.3460318